\documentclass[12pt]{article}
\usepackage{enumerate}
\usepackage[colorlinks=true,urlcolor=blue]{hyperref}

\begin{document}

\begin{small}

\begin{center}
{\bf MATH 2552: Differential Equations  \\
 January -  May 2017\\
TR 9:35 - 10:55 at  Howey L3
}\\
\href{http://people.math.gatech.edu/~sscott42/teach/2552spring17/math2552.html}{Course Webpage}\\
\href{http://people.math.gatech.edu/~sscott42/teach/2552spring17/math2552.html}{http://people.math.gatech.edu/~sscott42/teach/2552spring17/math2552.html}
\end{center}


\begin{tabular}{l l l l}
Lecturer & Shane Scott &  & \\
Email & scottsha@gatech.edu &  &  \\
Office & Skiles 252 &  &  \\
Hours & on course site  & & \\
& or by appointment&&
\end{tabular}

\vspace{.25cm}
\begin{tabular}{l l l l}
 &D1 & D2&  D3\\
Instructor & Pedro Maddens & Renyi Chen &  Renyi Chen\\
Room & Skiles 202 & Skiles 170 & Skiles 156 \\
Time & MW 9:05-9:55& MW 9:05-9:55 & MW 10:05-10:55
\end{tabular}

\begin{enumerate}[]

\item Welcome to \href{http://www.math.gatech.edu/course/math/2552}{Differential Calculus}! Attend all lectures.
Relevant text sections will be posted to the course webpage.
Read the text sections and work through text examples to prepare for lecture.
%Come prepared with your e-clicker and questions.
Participate. 
%If you haven't interupted me with a question, you're doing it wrong.

\item  \textbf{Course Objectives}: 
Students will learn to construct, analyze, and solve differential equation with applications from biology, physics, and engineering.
We will learn solutions to common equations, numerical techniques for estimating solutions, and equilibrium phenomena.

\item \textbf{Resources:}
\begin{enumerate}[$\cdot$]
 \item  The \href{http://people.math.gatech.edu/~sscott42/teach/2552spring17/math2552.html}{Course Webpage} will update to keep abreast with our progress and carry course materials.
		\item \href{https://t-square.gatech.edu/portal}{T-square} will host grades. 
        \item \href{https://piazza.com/gatech/fall2015/math1551g/home}{Piazza} will host online class Q\&A and announcements.
		
		\item \href{https://www.math.gatech.edu/academics/undergraduate/tutors-and-labs}{Mathlab} in Culc 280 offers free tutoring from recitation instructors. Check the link for summer hours.
\end{enumerate}

\item \textbf{Textbook:}
\href{http://www.wiley.com/WileyCDA/WileyTitle/productCd-EHEP003244.html}{James R. Brannan and William E. Boyce, Differential Equations (an introduction to modern methods and applications), 3rd edition, John Wiley \& Sons, 2015.}

\item \textbf{Grade Policy:}
Homework will be assigned and posted to the course website on a weekly basis.
Quizzes over homework problems will be given in recitation. 
The lowest two quiz grades will be dropped.
In calss participation extra credit may be given for clicker response questions in lecture and as recitation participation grades.
There will be 2 midterm exams and a cumulative final.
Only paper and pencil are allowed in any quiz.
On exams you may also bring one sheet of hand written notes.
Grades will be computed with the following weights
\begin{enumerate}
\item Quizzes and Participation $30\%$
\item Two Midterms $20\%$ each
\item Final Exam $ 30\%$
\end{enumerate}
Cutoff grades will be F=[0,60), D=[60,70), C=[70,80), B=[80,90), A=[90,100].
Make-up examinations will be given only in the event of a valid, documented excuse.
Exam regrade requests may be made in writing within one week of return.

\item
\textbf{Students with Disabilities and/or in need of Special Accommodations:} Georgia Tech complies with the regulations of the Americans with Disabilities Act of 1990 and offers accommodations to students with disabilities. If you are in need of Testing Accommodations and/or Note Taking Service, please contact the ADAPTS office.

\item 
{
\textbf{Honor Code:}
All students are expected to comply with the Georgia Tech Honor Code. Any
evidence of cheating or other violations of the Georgia Tech Honor Code will be
submitted directly to the Office of Student Integrity. The institute honor code is available
at \href{http://www.honor.gatech.edu}{http://www.honor.gatech.edu}.
}




\item \textbf{Important \href{http://www.registrar.gatech.edu/calendar/}{Dates}}:
\begin{flushleft}
First Class \dotfill 9 January \\
Add/Drop Deadline \dotfill 13 January\\
Exam 1 \dotfill 9 Feb\\
Withdraw Deadline \dotfill 15 March\\
Spring Break \dotfill 20-24 March\\
Exam 2 \dotfill 30 March\\
Last Class \dotfill 25 April\\
Course Final \dotfill Tuesday  2 May 2:50 - 5:40
\end{flushleft}

\end{enumerate}



\end{small}


\end{document}
